---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Prepare R Environment & load necessary library
```{r}
library(ggplot2)
library(dplyr)
```

The source data is loaded into a data frame 'data' for futher analysis
```{r}
data <- read.csv("activity.csv")
```

Explore the data frame to have a feel of the data
```{r}
str(data)
summary(data)
```

Convert Date Column to Date Format
```{r}
data$date <- as.Date(data$date)
```

## What is mean total number of steps taken per day?
Total number of steps per day
```{r}
StepsPerDay <- data %>% 
                group_by(date) %>% 
                summarise(TotalSteps=sum(steps))

hist(StepsPerDay$TotalSteps, breaks = 10,xlab='Total Steps',main='Histogram of Total Steps Per Day')
```

Mean & Median Number of Steps taken
```{r}
mean(StepsPerDay$TotalSteps, na.rm = TRUE)
median(StepsPerDay$TotalSteps, na.rm = TRUE)
```

## What is the average daily activity pattern?
Plot of Average Steps for each 5 minutes interval
```{r}
StepsPerInterval <- data %>% 
                    group_by(interval) %>%
                    summarise(AverageSteps=mean(steps,na.rm=T),TotalSteps=sum(steps,na.rm=T))

with(StepsPerInterval,plot(interval,AverageSteps, type = 'l', xlab='Interval', ylab='Average Steps', main='Average Steps for 5 Minutes Interval'))
```
  
Interval with highest number of steps
```{r}
StepsPerInterval[which.max(StepsPerInterval$TotalSteps),1]
```

## Imputing missing values
Total Rows with NA
```{r}
sum(is.na(data$steps))
```
  
Missing Values will be imputed with avarage steps of the same time interval (rounded to the nearest integer)
```{r}
Impute <- data %>% 
            inner_join(StepsPerInterval,by='interval') %>%
            mutate(NewSteps=ifelse(is.na(steps),round(AverageSteps),steps))
```

New Imputed dataset
```{r}
Impute <- Impute[,c(2,3,6)]
summary(Impute)
```

Total Steps per day (Imputed dataset)
```{r}
ImputePerDay <- Impute %>% 
                group_by(date) %>% 
                summarise(NewTotalSteps=sum(NewSteps))

hist(ImputePerDay$NewTotalSteps, breaks=10, xlab='Total Steps', main='Histogram of Total Steps Per Day (Imputed Dataset)')
```

Mean & Median Number of Steps taken (Imputed dataset)
```{r}
mean(ImputePerDay$NewTotalSteps, na.rm = TRUE)
median(ImputePerDay$NewTotalSteps, na.rm = TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?
Label the dates with Weekday & Weekend
```{r}
Impute$Days <- weekdays(Impute$date)
Impute$Days <- ifelse(Impute$Days=='Sunday'|Impute$Days=='Saturday','weekend','weekday')

AvgImpute <- Impute %>% 
                group_by(Days, interval) %>%
                summarise(TotalSteps=sum(NewSteps),AvgSteps=mean(NewSteps))
```

Time Series Plot
```{r}
ggplot(AvgImpute, aes(interval,AvgSteps)) + geom_line() + facet_wrap(~Days,2,1) + ggtitle("Avarage Steps Taken\n(For Weekday & Weekend)")
```
